---
title: "InClass_Quiz5_Meyer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

“All work presented is my own. I have not communicated with or worked with anyone else on this quiz.”

Trent Meyer

# Question 1: Consider the following two bar plots using the palmerpenguins data set. The first is a plot of the penguin species while the second is a plot of the average bill length for each species.
# Plot 1
```{r}
library(palmerpenguins)
library(tidyverse)
ggplot(data = penguins, aes(x = species)) +
  geom_bar() +
  labs(y = "Count")
```

# Plot 2
```{r}
ggplot(data = penguins %>% group_by(species) %>%
         summarise(avg_length = mean(bill_length_mm, na.rm = TRUE)),
       aes(x = species, y = avg_length)) +
  geom_col() +
  labs(y = "Average Bill Length")
```

Which of the two graphs is appropriate to construct? Give a one sentence reason.

Plot 1 is appropriate because we are not applying it a count, rather than a summary statistic. Plot 2 is plotting average bill length, a summary statistic, that is continuous, which could be better shown with a point plot, or another type of graph.

# Question 2: Use the Happy Planet Index data set to construct a graph that does not properly show variability in the underlying data. Recall that some variables in this data set are LifeExpectancy, Wellbeing, Footprint, and Region of the world.

```{r}
library(here)
hpi_df <- read_csv(here("data/hpi-tidy.csv"))

hpi_df_2 <- hpi_df %>% filter(HPIRank <= 10)

ggplot(data = hpi_df_2, aes(x = Country, y = LifeExpectancy)) +
  geom_col()

hpi_df_2 <- hpi_df_2 %>% mutate(Country = fct_reorder(Country, LifeExpectancy))


```

# Question 3: Fix your graph from the previous question so that it does properly show variability in the underlying data.

```{r}
ggplot(data = hpi_df_2, aes(x = Country, y = LifeExpectancy)) +
  geom_point() + 
  geom_segment( aes(x=Country, xend=Country, y=0, yend=LifeExpectancy)) +
coord_flip()
```

