---
title: "Quiz5_Meyer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

“All work presented is my own, and I have followed all rules for collaboration.”

Trent Meyer

```{r}
library(tidyverse)
statsurvey_df <- read_csv("data/stat113_survey.csv")
```

# Question 1: Construct a graph that uses the number of Facebook friends each student has.

```{r}
stat_question1 <- statsurvey_df %>% filter(!is.na(Facebook)) %>%
  group_by(time_both) %>%
  summarise(meanfacebook = mean(Facebook),
            sd = sd(Facebook),
            count = n()) %>%
  mutate(se = sd / sqrt(count),
         u_se = meanfacebook + se,
         l_se = meanfacebook - se)

stat_question1 <- stat_question1 %>% separate(time_both, into = c("semester", "year"),
                               sep = 1) %>%
  arrange(year, desc(semester)) %>%
  unite(col = "time_both", c(semester, year)) %>%
  mutate(time_both = fct_inorder(time_both))


ggplot(data = stat_question1, aes(x = time_both, y = meanfacebook)) +
  geom_point() +
  geom_errorbar(aes(ymin = l_se, ymax = u_se)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
```

# Question 2: Construct a graph that uses the proportion of students that have Facebook (assuming that 0 friends means that the student does not have a Facebook account and more than 0 friends means that the student does have a Facebook account).

```{r}
stat_question2 <- statsurvey_df %>% filter(!is.na(Facebook)) %>%
  mutate(iffacebook = if_else(Facebook == 0,
                              "No",
                              "Yes")) 

stat_question2 <- stat_question2 %>% separate(time_both, into = c("semester", "year"),
                               sep = 1) %>%
  arrange(year, desc(semester)) %>%
  unite(col = "time_both", c(semester, year)) %>%
  mutate(time_both = fct_inorder(time_both))

ggplot(data = stat_question2, aes(x = time_both, fill = iffacebook)) +
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

