---
title: "Quiz4_Meyer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

“All work presented is my own, and I have followed all rules for collaboration.”

Trent Meyer

Use the leaflet package and the us.cities data set from the maps package to make a Zoom-able map of the 50 U.S. capital cities. When you click on a marker for a capital city, both the city name and state as well as the pop (population) should be displayed.

Note that, in the us.cities data set, state capitals are given a value of 2 in the capital variable.

A second note is that if you cannot figure out how to get two “things” (in this case name and population) to display in Leaflet, you would just lose a single point for the quiz (so don’t spend too much time on that part unless you are inclined to do so).

```{r}
library(leaflet)
library(maps)
library(tidyverse)
uscities <- maps::us.cities
uscities <- uscities %>% filter(capital == 2)

content <- uscities %>%
  mutate(popup = paste0('</a>', uscities$name, ' ',
                        uscities$pop, ' </a>'))

cities_map <- leaflet(uscities) %>%
  setView(lng = -98.583, lat = 39.833, zoom = 4) %>% 
  addTiles() %>% 
  addProviderTiles(providers$Wikimedia) %>% 
  addMarkers(lng = uscities$long, lat = uscities$lat,
             clusterOptions = markerClusterOptions(),
             popup = content$popup)
cities_map
```

