---
title: "Quiz02_Meyer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

“All work presented is my own, and I have followed all rules for collaboration.”

Trent Meyer

```{r}
library(tidyverse)
library(readxl)
df <- read_excel("data/slu_graduates_17_21.xlsx")

## fixes error in the data
df <- df %>% mutate(across(everything(),
                           .fns = ~replace(., . ==  "STATS" , "STAT")))
```

# Question 1: Make the following chart that shows the “other” majors of each STAT major student in the data set. 

```{r}
quiz_df <- df %>% filter(major1 == "STAT" | major2 == "STAT" | major3 == "STAT") %>%
  pivot_longer(c(3:8), 
               names_to = "type",
               values_to = "discipline") %>%
  filter(type == "major1" | type == "major2" | type == "major3") %>%
  filter(!is.na(discipline) & discipline != "STAT") %>%
  group_by(discipline) %>%
  summarise(nstudents = n()) %>%
  arrange(desc(nstudents)) %>%
  mutate(discipline = fct_reorder(discipline, nstudents))
  
df_stat <- df %>% pivot_longer(c(3:8),
                    values_to = "discipline",
                    names_to = "type") %>%
  filter(discipline == "STAT")
semi_join(df, df_stat, by = "adm_id")

ggplot(data = quiz_df, aes(x = discipline, y = nstudents)) +
  geom_col() + 
  coord_flip()
```

# Question 2: Make the following lollipop graphic, which investigates the sex disparity of majors at SLU with 50 or more total graduates in the past few years. The y-axis is the major while the x-axis is the proportion of majors who identified as Female. Note that the graph shown below eliminates some of the majors because I want you to discover the final trends on your own! Remember that your graphic should only include all of the majors with 50 or more total graduates.

```{r}
question2_df <- df %>% select (1:8) %>%
  pivot_longer(c(3:8),
                    values_to = "discipline",
                    names_to = "type") %>%
  filter(type == "major1" | type == "major2" | type == "major3") %>%
  filter(!is.na(discipline)) %>% 
  group_by(sex, discipline) %>%
  summarise(nstudents = n()) %>%
  ungroup() %>%
  group_by(discipline) %>%
  mutate(ntotal = sum(nstudents),
         femaleprop = nstudents / ntotal) %>%
  filter(nstudents >= 50,
         sex == "F") %>%
  ungroup() %>%
  arrange(desc(femaleprop)) %>%
  mutate(discipline = fct_reorder(discipline, femaleprop))
  

ggplot(data = question2_df, aes(x = discipline, y = femaleprop)) +
  geom_point() +
  geom_segment(aes(x=discipline, xend=discipline, y=0, yend=femaleprop)) +
  coord_flip() +
  labs(x = "discipline",
       y = "Proportion of Female Majors in Past Five Years")
```

